Non-kissing and non-crossing complexes \\ for locally gentle algebras

### 2018-07-09 draft/accordionComplexSurfaces.tex

@@

\begin{proof}
Let~$c=\alpha_1\cdots\alpha_r$ be a non-trivial oriented cycle in some walk~$\omega$, with~$c, c^2\notin I$.
Assume that~$\omega$ is not of form~$\sigma c^{\pm\infty}$.
Then, up to cyclic reordering of the arrows in~$c$, it contains a substring of the form~$\beta^\pm c^l \alpha_1\cdots\alpha_j\gamma^\pm$, for some arrows~$\beta,\gamma$ not in~$c$, some~$l\geq 1$, and some~$j<r$.
Since~$\bar Q$ is locally gentle, this substring is~$\beta^{-1} c^l \alpha_1\cdots\alpha_j\gamma^{-1}$.
Hence, the walk~$\omega$ contains the two substrings~$\beta^{-1}\alpha_1\cdots\alpha_j\alpha_{j+1}$ and~$\alpha_r\alpha_1\cdots\alpha_j\gamma^{-1}$, and thus self-kisses.
\vincent{Should we also give here a geometric argument?}
\end{proof}

@@

\begin{proposition}
\label{prop:purity}
The non-kissing complex~$\NKC$ and the non-crossing complex~$\NCC[\dissection_{\bar Q}, {\dissection_{\bar Q}}\dual]$ are pure of dimension~$3|Q_0| - |Q_1|$.
\vincent{This is wrong. We need to count the infinite straight walks. I propose the next statement instead.}
\end{proposition}

@@

\para{Geometric proof}
\vincent{Should we also give here a geometric argument? Pierre-Guy, could you add that?}

>>>>

The curves~$\curveof(\omega)$ and~$\curveof(\omega')$ on the surface follow~$\omega$ and~$\omega'$ on the picture.
It is then clear that~$\omega$ and~$\omega'$ kiss if and only if~$\curveof(\omega)$ and~$\curveof(\omega')$ cross. 
\end{proof}

@@

\begin{proof}
Let~$c=\alpha_1\cdots\alpha_r$ be a non-trivial oriented cycle in some walk~$\omega$, with~$c, c^2\notin I$.
Assume that~$\omega$ is not of form~$\sigma c^{\pm\infty}$.
Then, up to cyclic reordering of the arrows in~$c$, it contains a substring of the form~$\beta^\pm c^l \alpha_1\cdots\alpha_j\gamma^\pm$, for some arrows~$\beta,\gamma$ not in~$c$, some~$l\geq 1$, and some~$j<r$.
Since~$\bar Q$ is locally gentle, this substring is~$\beta^{-1} c^l \alpha_1\cdots\alpha_j\gamma^{-1}$.
Hence, the walk~$\omega$ contains the two substrings~$\beta^{-1}\alpha_1\cdots\alpha_j\alpha_{j+1}$ and~$\alpha_r\alpha_1\cdots\alpha_j\gamma^{-1}$, and thus self-kisses.

\para{Geometric proof}
An oriented cycle~$c$ corresponds to a puncture in~$\dissection_{\bar Q}$.
Let~$\gamma = \curveof(\omega)$.  The fact that~$\omega$ contains~$c$ translates to the fact that~$\gamma$ makes a full circle around the correponding puncture.
\pierreguy{Pic No 8 here.}

\includegraphics{pgFigNo8.pdf}

Once this cycle is completed, since~$\gamma$ cannot cross itself, it has to continue cycling infinitely around the puncture.  This proves the claim.
\vincent{Should we also give here a geometric argument?}
\end{proof}

@@

@@

\para{Geometric proof}
The statement of the lemma translates on~$\surface(\bar Q)$ as follows: for any facet~$F$ of the non-crossing complex~$\NCC$ and any puncture of the dissection~$\dissection_{\bar Q}$, there exists a curve in~$F$ which goes to the puncture.
Assume that there is a puncture with no curves of~$F$ going to it.
Construct a curve~$\gamma$ as follows:~$\gamma$ first unwinds from the puncture, then~$\gamma$ leaves the local picture around the puncture by turning aroune another marked point (marked~$X$ on the picture below).

\includegraphics{pgFigNo7.pdf}
\pierreguy{Pic No 7 here.}
Two situations occur.

Firstly, if there are no curves of~$F$ closer to~$X$ than~$\gamma$, then~$\gamma$ can circle around~$X$ until it reaches the boundary (if~$X$ is on the boundary) or infinitely (if~$X$ is a puncture).
Then~$\gamma$ is a~$\dissection_{\bar Q}$-accordion which crosses none of those in~$F$, contradicting the maximality of~$F$.

Secondly, if there is a curve of~$F$ closer to~$X$ than~$\gamma$, then let~$\delta$ be the one furthest from~$X$ ($\delta$ exists by Lemma \ref{lemm:facesHaveMaximalElement}).
From there, let~$\gamma$ follow~$\delta$ on the surface until it either reaches the boundary or circles infinitely around a puncture.  
Then~$\gamma$ is compatible with all curves in~$F$, a contradiction.

Thus~$F$ contains a curve going to the puncture.
\vincent{Should we also give here a geometric argument? Pierre-Guy, could you add that?}
\pierreguy{Done; I will need a picture, please, Vincent.}
